\chapter{Acknowledgements}

Firstly, I would like to express my deep and sincere gratitude to my adviser Professor Dag Haugland for his guidance and encouragement during the four years I spent as his PhD student.
Without his constant feedback and immense patience, this thesis and the publications written throughout my doctoral education would not have been attainable.
I learned a lot from my adviser, and gained valuable experience for which I am deeply indebted.
His hardworking, passionate, and precise attitude inspired me greatly in my academic growth.

I am also very grateful to doc. RNDr. Pavel Surynek, PhD, from Czech Technical University, 
for giving me the precious and unforgettable opportunity to spend three months as an intern in Artificial Intelligence Research Center in Japan.
As a co-author of two papers resulting from my internship, he contributed to this thesis as well. 

My sincere thanks goes to the entire Department of Informatics.
It was a pleasure to work in such an inspiring and accepting environment. 
In particular, I would like to mention Professor Jan-Joachim R\"uckmann for his advice and affability, 
and my fellow doctoral students from the Optimization group for being good friends and making my days enjoyable and exciting.
In addition, I greatly appreciate the work of administration staff for helping me with any organisational issue.

I must not forget to thank my husband Jahan Zeb for his endless patience and support during the ups and downs of the entire PhD.
Last but not least my parents have always believed in me, encouraged and motivated me in my studies. 
Anything that I have achieved in my life would be impossible without their unconditional support. 

